\Title

### 2023-01-19 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

%\ninept
%
%\abstract{Abstract}
%
\keywords{\Keywords}
%
\maketitle
%%%%%%%%% BODY TEXT
\section{Introduction}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
\label{sec:intro}
% >>>>>>>>>>> LuP is here <<<<<<<<<<<
%
\subsection{Recent challenges in machine learning and computer vision}
%%%-----------------------------------------------------------------
Machine learning, and computer vision in particular, have undergone a major transformation in modern times, transitioning from an eccentric character in science fiction movies to become mainstream tools found in the smartphones in everyone's pocket. Less visible, yet crucial to our economy and security, these automated tools now dominate the range of services that equip virtually most sectors of our daily lives. This trend is profoundly pronounced in the field of medicine, as it provides practitioners with new diagnostic tools to treat people faster and more effectively. This particular example also illustrates some important challenges that these tools are increasingly facing. One of these is the sensitivity of deep-learning based algorithms to adversarial attacks, that is, the sensitivity of the models' output to minute changes in the input. This could raise important questions about the reliability of the results of any automated diagnostic model. More generally, these models do give an output for any possible input, but they lack an explanation as to why they produced it. This lack of explainability is a hindrance to the advancement of these methodologies, as well as to their acceptance for most safety-relevant tasks. Having said that, we may remember that these tools are inherited from theoretical models originally designed to understand the brain. While the first artificial neural networks were mainly applied to toy-like problems, they also offered a growing number of applications in machine learning and computer vision, but also more generally for task automation. In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates, and today they reach higher performances than humans for key image categorization challenges. One of the challenges of these models is to be able to scale them up for future applications, for example in autonomous driving. Indeed, in such applications, the input does not consist of a single image (say a few million pixels) but of a continuous stream of images from several cameras.

The challenges are numerous in such situations, and two major obstacles must be highlighted: the computational processing time and the energy budget. First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun_gradient-based_1998}, then high definition images, such as VGG on ImageNet~\citep{simonyan_very_2015}. This has a cost and notably the fact that the computation time increases with the size (in number of pixels) of the image. This phenomenon has been mitigated by the increase in available computing power, notably thanks to the parallel processing implemented in GPU cards. However, the transition from images to videos, then to multiple cameras implies a further level in the change of scale in the amount of data processed. It also raises an issue that is often not addressed when it comes to image categorization, that is, of the right timing to take a decision. Indeed, in autonomous driving, an important task is to react as quickly as possible to a possible collision, for example when a pedestrian unexpectedly crosses the street. As such, future systems should always be able to give a response at any time, instead of having to wait for the pipeline to provide the correct answer. This online processing is also a feature of processing in biological systems and could be an important first contribution if we could translate these principles into future automated processes. %
Second, a major constraint for future systems is energy cost. As we have seen, fast computation time has been maintained for high-dimensional images, but at the cost of higher energy consumption. In addition, the need to solve increasingly complex tasks has required the construction of increasingly complex and deep architectures. As an example, the best performing model in the ImageNet challenge uses 2100 million parameters. Training the model typically consumes many kilowatts hours of energy and about 1 joule to infer a decision per single image. Less power-hungry solutions have emerged, such as Mobile Net~\citep{howard_searching_2019}, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significant percentage of the energy consumed by an electric car under standard conditions.

@@

Neuroscience could provide revolutionary answers to these two closely related problems. Indeed, the human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, about tens of watts. This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use. CNNs, for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels. These networks have been known to mimic many properties of biological systems~\citep{schrimpf_brain-score_2020}, yet, this score does not count rapidity or energy. On the other hand, information in a large majority of biological neural networks is represented as spikes, i.e. prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike~\citep{paugam-moisy_computing_2012}. Why spikes? It is not yet known why most biological systems have evolved to use this representation, but one major advantage seems to be some form of efficiency. In the human brain, each of the approximately 90 trillion nerve cells emits on average one spike per second which fan out to an average of 10000 synapses. Importantly, each neuron fires with great variability, and up to 300 spikes per second. Many attempts have been made to build artificial spiking neural networks (SNNs), and in particular to achieve the same performance as observed in classical CNNs with a lower energy budget. However, none of these networks has yet surpassed the state-of-the-art performances. Here, we will try to focus on one core property of spiking neurons, inspired by neurobiology to progress in this direction.

>>>>

%\ninept
%
%\abstract{Abstract}
%
\keywords{\Keywords}
%
\maketitle
%%%%%%%%% BODY TEXT
\section{Introduction}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
\label{sec:intro}
% >>>>>>>>>>> LuP is here <<<<<<<<<<<
%
\subsection{Recent challenges in machine learning and computer vision}

The human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, about tens of watts. This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use. Convolutional Neural Networks (CNNs), for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels. These networks have been known to mimic many properties of biological systems~\citep{schrimpf_brain-score_2020}, yet, this score does not account for rapidity or energy consumption. On the other hand, information in a large majority of biological neural networks is represented as \emph{spikes}, i.e. prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike~\citep{paugam-moisy_computing_2012}. While CNNs obtain state-of-the-art performances for some computer vision tasks like image recognition, they do not make use of the dynamics inherent to the way we sense our natural environment. The third generation of artificial neural networks, i.e. Spiking Neural Networks (SNNs), include this temporal dimension in their computations. This new kind of algorithms are interesting for computational neuroscience as they are a better model of biological neurons. Another class of approaches have designed normative models of SNNs that aim to have applications in machine learning. One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}. Another type of SNN using precise spike timing attempts to determine the structure of the network in order to minimize a cost function. This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extended in novel gradient-based methods. The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}. However, the performance of SNNs is still lagging that of firing rate-based networks and the question of the advantage of using spikes in machine learning and computer vision remains open.

%%%-----------------------------------------------------------------
%Machine learning, and computer vision in particular, have undergone a major transformation in modern times, transitioning from an eccentric character in science fiction movies to become mainstream tools found in the smartphones in everyone's pocket. Less visible, yet crucial to our economy and security, these automated tools now dominate the range of services that equip virtually most sectors of our daily lives. This trend is profoundly pronounced in the field of medicine, as it provides practitioners with new diagnostic tools to treat people faster and more effectively. This particular example also illustrates some important challenges that these tools are increasingly facing. One of these is the sensitivity of deep-learning based algorithms to adversarial attacks, that is, the sensitivity of the models' output to minute changes in the input. This could raise important questions about the reliability of the results of any automated diagnostic model. More generally, these models do give an output for any possible input, but they lack an explanation as to why they produced it. This lack of explainability is a hindrance to the advancement of these methodologies, as well as to their acceptance for most safety-relevant tasks. Having said that, we may remember that these tools are inherited from theoretical models originally designed to understand the brain. While the first artificial neural networks were mainly applied to toy-like problems, they also offered a growing number of applications in machine learning and computer vision, but also more generally for task automation. In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates, and today they reach higher performances than humans for key image categorization challenges. One of the challenges of these models is to be able to scale them up for future applications, for example in autonomous driving. Indeed, in such applications, the input does not consist of a single image (say a few million pixels) but of a continuous stream of images from several cameras.

%The challenges are numerous in such situations, and two major obstacles must be highlighted: the computational processing time and the energy budget. First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun_gradient-based_1998}, then high definition images, such as VGG on ImageNet~\citep{simonyan_very_2015}. This has a cost and notably the fact that the computation time increases with the size (in number of pixels) of the image. This phenomenon has been mitigated by the increase in available computing power, notably thanks to the parallel processing implemented in GPU cards. However, the transition from images to videos, then to multiple cameras implies a further level in the change of scale in the amount of data processed. It also raises an issue that is often not addressed when it comes to image categorization, that is, of the right timing to take a decision. Indeed, in autonomous driving, an important task is to react as quickly as possible to a possible collision, for example when a pedestrian unexpectedly crosses the street. As such, future systems should always be able to give a response at any time, instead of having to wait for the pipeline to provide the correct answer. This online processing is also a feature of processing in biological systems and could be an important first contribution if we could translate these principles into future automated processes. %
%Second, a major constraint for future systems is energy cost. As we have seen, fast computation time has been maintained for high-dimensional images, but at the cost of higher energy consumption. In addition, the need to solve increasingly complex tasks has required the construction of increasingly complex and deep architectures. As an example, the best performing model in the ImageNet challenge uses 2100 million parameters. Training the model typically consumes many kilowatts hours of energy and about 1 joule to infer a decision per single image. Less power-hungry solutions have emerged, such as Mobile Net~\citep{howard_searching_2019}, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significant percentage of the energy consumed by an electric car under standard conditions.

@@

%Neuroscience could provide revolutionary answers to these two closely related problems. Indeed, the human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, about tens of watts. This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use. CNNs, for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels. These networks have been known to mimic many properties of biological systems~\citep{schrimpf_brain-score_2020}, yet, this score does not count rapidity or energy. On the other hand, information in a large majority of biological neural networks is represented as spikes, i.e. prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike~\citep{paugam-moisy_computing_2012}. Why spikes? It is not yet known why most biological systems have evolved to use this representation, but one major advantage seems to be some form of efficiency. In the human brain, each of the approximately 90 trillion nerve cells emits on average one spike per second which fan out to an average of 10000 synapses. Importantly, each neuron fires with great variability, and up to 300 spikes per second. Many attempts have been made to build artificial spiking neural networks (SNNs), and in particular to achieve the same performance as observed in classical CNNs with a lower energy budget. However, none of these networks has yet surpassed the state-of-the-art performances. Here, we will try to focus on one core property of spiking neurons, inspired by neurobiology to progress in this direction.
